[Artifact Presentation] Damegender: \\ Writing and Comparing Gender Detection Tools

### 2020-05-15 paper2sattose.tex
paper2sattose.pdf, paper2sattose.tex: new revision

In Section 2, we explain Damegender as solution to the problem,
explaining that we are using Open Datasets extracted from the states
as base of truth. Section 3, presents a feature comparison with other
tools.  Section 4, reproduces accuracies and confusion matrix with a
scientific dataset. The section 5 is about how we are using Machine
Learning in Damegender. The section 6 is about conclusion, limitations
and futher research.

@@

\subsection{Datasets}

@@

In \texttt{damegender}, we are including Open Data census about names and gender, from institutions such as INE.es (the Spanish National Statistics Institute), or the governments of Uruguay, USA and United Kingdom.
%We want datasets provided by the software package to increment the speed retrieving data.
Some Open Datasets, such the one offered by INE.es or the government of the United States of America offer support for surnames and how they are related to ethnicity. 
In particular, the dataset from the government of the United States of America offers a probability of the race, and the Spanish INE.es gives the number of people with a surname with a nationality different to the Spanish nationality.

@@

\begin{table*}[ht]
\footnotesize
\begin{tabular}[]{lcccccc}
\hline
Service / Tool $->$ & Gender API & gender-guesser & genderize.io & NameAPI & NamSor & damegender\tabularnewline
\hline
Database size & 431M & 45K & 114M & 1M & 4G & 57K \tabularnewline
Regular data updates & Yes & no & No & Yes & Yes & Yes\tabularnewline
Unstructured full name strings & Yes & No & No & Yes & No & Yes\tabularnewline
Surnames & Yes & No & No & Yes & Yes & Yes\tabularnewline
Non-Latin alphabets & Partial & No & Partial & Yes & Yes & No\tabularnewline
Implicit geo-localization & Yes & No & No & Yes & Yes & No\tabularnewline
Exists locale & Yes & Yes & Yes & Yes & Yes & Yes\tabularnewline
Assingment type & P & B & P & P & P & P \tabularnewline
Free parameters & T,P & G & P,C & T & S & T,C\tabularnewline
Prediction & No & No & No & No & No & Yes\tabularnewline
Free license & No & Yes & No & No & No & Yes\tabularnewline
REST API & Yes & No & Yes & Yes & Yes & Planned\tabularnewline
Limits number of requests & Yes (200) & \infinity & Yes & Yes & Yes & \infinity \tabularnewline
\hline
\end{tabular}
\caption{Comparison of the different features that gender guesser software services and tools offer. Assignment type = \{P: Probabilistic; B: Binary\}. Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \}. \grex{Estaria bien contar con ``the monthly subscription cost (calculated for 100,000 requests/month)'' en la tabla, porque daría más info muy relevante.} }
\label{table:comparison}
\end{table*}

@@

Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} offer a framework to classify gender tools.
The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).

@@

Santamara and Mihaljevi~\cite{10.7717/peerj-cs.156} explain different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
In their dataset, gender is classified as male, female or unknown. 
We have used this dataset, not considering the unknow variable, for our experiments.
The results can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
As can be observed, Genderapi and Genderize are obtaining the best results, although all solutions reach results better than 0.8, except for \texttt{Gender Guesser}.

\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}

@@

In Table~\ref{table:ApisAndErrors} we can observe a high index of errors in \texttt{Nameapi} and \texttt{Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.

\grex{Aqui haria falta indicar por que hacemos esta comparacion y que queremos aprender de ella}

@@

Here the acknowledgements come...

>>>>

In Section 2, we explain Damegender as solution to the
problem. Section 3 explains that we are using Open Datasets extracted
from the states as base of truth. Section 4, presents a feature
comparison with other tools.  Section 5, reproduces accuracies and
confusion matrix with a scientific dataset. The section 6 is about how
we are using Machine Learning in Damegender. The section 7 is about
conclusion, limitations and futher research.

@@

\section{Datasets}

@@

In \texttt{damegender}, we are including Open Data census about names and gender, from institutions such as INE.es (the Spanish National Statistics Institute), or the governments of Uruguay, USA and United Kingdom. The datasets provided by the software package is incrementing the speed retrieving data.

Some Open Datasets, such the one offered by INE.es or the government of the United States of America offer support for surnames and how they are related to ethnicity. 
In particular, the dataset from the government of the United States of America offers a probability of the race, and the Spanish INE.es gives the number of people with a surname with a nationality different to the Spanish nationality.

@@

\begin{table*}[ht]
\footnotesize
\begin{tabular}[]{lcccccc}
\hline
Service / Tool $->$ & Gender API & gender-guesser & genderize.io & NameAPI & NamSor & damegender\tabularnewline
\hline
Database size & 431M & 45K & 114M & 1M & 4G & 57K \tabularnewline
Regular data updates & Yes & no & No & Yes & Yes & Yes\tabularnewline
Unstructured full name strings & Yes & No & No & Yes & No & Yes\tabularnewline
Surnames & Yes & No & No & Yes & Yes & Yes\tabularnewline
Non-Latin alphabets & Partial & No & Partial & Yes & Yes & No\tabularnewline
Implicit geo-localization & Yes & No & No & Yes & Yes & No\tabularnewline
Exists locale & Yes & Yes & Yes & Yes & Yes & Yes\tabularnewline
Assingment type & P & B & P & P & P & P \tabularnewline
Free parameters & T,P & G & P,C & T & S & T,C\tabularnewline
Prediction & No & No & No & No & No & Yes\tabularnewline
Free license & No & Yes & No & No & No & Yes\tabularnewline
REST API & Yes & No & Yes & Yes & Yes & Planned\tabularnewline
Limits number of requests & Yes (200) & \infinity & Yes & Yes & Yes & \infinity \tabularnewline
\hline
\end{tabular}
\caption{Comparison of the different features that gender guesser software services and tools offer. Assignment type = \{P: Probabilistic; B: Binary\}. Free Parameters = \{T: total\_names; P: probability; C: count; G: gender; T: trust; S: scale \}. }
\label{table:comparison}
\end{table*}

\grex{Estaria bien contar con ``the monthly subscription cost (calculated for 100,000 requests/month)'' en la tabla, porque daría más info muy relevante.}
\davidam{Si te da tiempo por mi adelante}

@@

Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} offers a framework to classify gender tools.
The features observed in this framework are: (i) database size (as of January 2018), (ii) if there are regular data updates, (iii) if they handle unstructured full name strings, (iv) if they handle surnames, (v) if they handle non-Latin alphabets, (vi) if implicit geolocalization is available, (vii) if locale exists, (viii) the type of assignment, (ix) if free parameters are possible, (x) if they offer prediction, (xi) if the tool is released under an open source license, (xii) if they offer an API, (xiii) the amount of monthly free requests, and (xiv) the monthly subscription cost (calculated for 100,000 requests/month)).

@@

Santamaria and Mihaljevic~\cite{10.7717/peerj-cs.156} explains different ways to determine gender from a name by humans and offers 7,000 names applying these methods. 
In their dataset, gender is classified as male, female or unknown. 
We have used this dataset, not considering the unknow variable, for our experiments.
The results can be seen in Table~\ref{table:DifferentAccuraciesMeasures}.
As can be observed, Genderapi and Genderize are obtaining the best results, although all solutions reach results better than 0.8, except for \texttt{Gender Guesser}.

\grex{Deberiamos explicar que es cada una de las metricas. Y tambien por que tenemos 1.0 de recall en todos los casos.}
\davidam{Esto es un poco largo de explicar. Está explicado de una manera abreviada más arriba. Si te ves con energías para explicarlo, está explicado en https://easychair.org/publications/preprint/vthL}

@@

In Table~\ref{table:ApisAndErrors} we can observe the different measures for errors in the APIs. The most visible conclusion is an high index of errors in \texttt{Nameapi} and \texttt{Namsor} and a low index of errors in \texttt{GenderApi} and \texttt{damegender}.

@@

Thanks to:
Gregorio Robles by the support and suggestions for publish this article in Sattose

Lucía Santamaría and Helena Mihaljevic by the previous work.

Daniel Izquierdo and Laura Arjona for starts this research field in the URJC.

Jesus Gonzalez Boticario and the people who is working with him in the UNED by the motivation towards the machine learning.
